### 2021-11-24 Complexity/UGC/approx.tex
Grammar and additions to pcp section

\section{Approximation Algorithms and CSPs}
%3.2 Limits of Approximation – Introduction (TIFR)
%AB - Chapter 11
%Explain intuition.
\subsection{Definitions and Examples}
The theory of approximating $\NP$-hard problems roots itself in the following question: ``Is it possible to efficiently approximate $\NP$-complete problem to some arbitrary degree of accuracy?" Since Cook-Levin result demonstrated that the $\SAT$ decision problem is $\NP$-complete, the question could be rephrased as ``Can we find an efficient algorithhm for $\SAT$ which outputs an assignment which satisfies a $1-\delta$ fraction of the clauses for any constant $\delta >  0$?" To this end, let us first introduce a few motivating examples which will find utility in our analysis. \newline

@@

Naturally, $\mathsf{Max3CNF}$ is $\NP$-hard since its corresponding decision problem $\mathsf{3CNF}$ is $\NP$-complete i.e $\phi$ is satisfiable iff $\mathsf{Opt}(\phi) = 1$.
%
However, we can ask if there exists an algorithm $A$ which outputs an assignment which satisfies at least some $\beta\cdot\mathsf{Opt}(\phi)$ fraction of the clauses of $\phi$ for some $\beta \leq 1$. To formalize this: \newline
%
\begin{definition} \label{Opt3SATDef}
For $\beta \leq 1$, a polynomial-time algorithm $A$ is deemed as a $\beta$-approximation algorithm for $\mathsf{Max3CNF}$ if $A(\phi)$ outputs an assignment which satisifies at least $\beta\cdot\mathsf{Opt}(\phi)$ fraction of $\phi$'s clauses for every 3CNF instance $\phi$. More specifically, the algorithm $A$ outputs some value such that:
\begin{equation}
  \beta\cdot\mathsf{Opt}(\phi) \leq A(\phi) \leq \mathsf{Opt}(\phi)
\end{equation}
If $\mathsf{Opt}(\phi) = 1$, then $\phi$ is said to be \emph{satisfiable}.
\end{definition}

@@

\begin{align}
  &  \min \sum_{(u,v) \in E} e_{u,v} \\
  & e_{u,v} \geq
  \max \begin{cases}
      x_1 + x_2 \\
      2 - (x_1 + x_2)
  \end{cases} \\
  & e_{u,v} \in [1,2] \\
  & x_v \in [0,1]
\end{align}
%
where the $e_{u,v}$ variables represent the edges $e \in E$ of the input graph and the $x_v$ represent the vertex variables $v \in V$. This will actually be a relaxation from its respective integer program in which the variables can take the integer values:

\begin{align}
  e_{u,v} = \begin{cases}
              1 \text{ if } e_{u,v} \in \mathcal{C} \\
              2 \text{ otherwise }
            \end{cases}
  x_v = \begin{cases}
              0 \text{ if } v \text{ is in partition } S \\
              1 \text{ otherwise }
        \end{cases}
\end{align}

@@

\begin{definition} \label{gapDef}
  Given a $\MaxCSP$ problem, we define $\GapMaxCSP{\alpha}{\beta}$ for $\alpha < \beta$ as the promise problem: given an instance $\I$:
  \begin{align}
      \I \in \mathsf{YES} & \iff \mathsf{Opt}(\I) \geq \beta \\
        \I \in \mathsf{NO} & \iff \mathsf{Opt}(\I) < \alpha
  \end{align}
\end{definition}

>>>>

\section{Approximation Algorithms and CSPs}
%3.2 Limits of Approximation – Introduction (TIFR)
%AB - Chapter 11
%Explain intuition.
\subsection{Definitions and Examples}
The theory of approximating $\NP$-hard problems roots itself in the following question: ``Is it possible to efficiently approximate $\NP$-complete problem to some arbitrary degree of accuracy?" Since Cook-Levin result demonstrated that the $\SAT$ decision problem is $\NP$-complete, the question could be rephrased as ``Can we find an efficient algorithhm for $\SAT$ which outputs an assignment which satisfies a $1-\delta$ fraction of the clauses for any constant $\delta >  0$?" To this end, let us first introduce a few motivating examples which will find utility in our forthcoming analysis. \newline

@@

An polynomial-time algorithm which solves $\MaxCNF$ is one which takes a 3CNF $\phi$ as input and outputs $\Opt(\phi)$. Naturally, $\mathsf{Max3CNF}$ is $\NP$-hard since its corresponding decision problem $\mathsf{3CNF}$ is $\NP$-complete i.e $\phi$ is satisfiable iff $\mathsf{Opt}(\phi) = 1$.
%
However, we can ask if there exists an algorithm $A$ which outputs an assignment which satisfies at least some $\beta\cdot\mathsf{Opt}(\phi)$ fraction of the clauses of $\phi$ for some $\beta \leq 1$. To formalize this: \newline
%
\begin{definition} \label{Opt3SATDef}
For $\beta \leq 1$, a polynomial-time algorithm $A$ is deemed as a $\beta$-approximation algorithm for $\mathsf{Max3CNF}$ if $A(\phi)$ outputs an assignment which satisifies at least $\beta\cdot\mathsf{Opt}(\phi)$ fraction of $\phi$'s clauses for every 3CNF instance $\phi$. More specifically, the algorithm $A$ outputs some value such that:
\begin{equation}
  \beta\cdot\mathsf{Opt}(\phi) \leq A(\phi) \leq \mathsf{Opt}(\phi)
\end{equation}
If $\mathsf{Opt}(\phi) = 1$, then $\phi$ is said to be \emph{satisfiable}.
\end{definition}

@@

\begin{align}
  &  \max \sum_{(u,v) \in E} e_{u,v} \\
  & e_{u,v} \leq
  \min \begin{cases}
      x_1 + x_2 \\
      2 - (x_1 + x_2)
  \end{cases} \\
  & e_{u,v} \in [0,1] \\
  & x_v \in [0,1]
\end{align}
%
where the $e_{u,v}$ variables represent the edges $e \in E$ of the input graph and the $x_v$ represent the vertex variables $v \in V$. This will actually be a relaxation from its respective integer program in which the variables can take the integer values:

\begin{align}
  e_{u,v} = \begin{cases}
              1 \text{ if } e_{u,v} \in \mathcal{C} \\
              0 \text{ otherwise }
            \end{cases}
  x_v = \begin{cases}
              0 \text{ if } v \text{ is in partition } S \\
              1 \text{ otherwise }
        \end{cases}
\end{align}

@@

\begin{definition} \label{gapDef}
  Given a $\MaxCSP$ problem, we define $\GapMaxCSP{\alpha}{\beta}$ for $\alpha < \beta$ as the promise problem: given an instance $\I$:
  \begin{align}
      \I \in \mathsf{YES} & \iff \mathsf{Opt}(\I) \geq \beta \\
        \I \in \mathsf{NO} & \iff \mathsf{Opt}(\I) < \alpha
  \end{align}
  Furthermore, an algorithm $A$ decides $\MaxCSP$ if for input instance $\I$ it accepts iff $\I \in \mathsf{Yes}$ and rejects iff $\I \in \mathsf{No}$. If $\I$ is in neither $\mathsf{Yes}, \mathsf{No}$, we do not care what the algorithm outputs.
\end{definition}

In particular, we deem a $\GapMaxCSP{\alpha}{\beta}$ problem as $\NP$-hard if for every language $L \in \NP$, there exists a polynomial-time reduction $f$ taking $x$ to CSP $\Psi$ instances such that:
\begin{align*}
  x \in L \implies \mathsf{Opt}(f(x)) & \geq \beta \\
  x \not\in L \implies \mathsf{Opt}(f(x))&  < \alpha
\end{align*}

### 2021-11-24 Complexity/UGC/pcp.tex
Grammar and additions to pcp section

\begin{theorem} \label{pcpgapsat}($\PCP$ Theorem: $\mathsf{Gap3SAT}$-hardness)
The problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max3SAT}$ is $\NP$-hard. In other words, for every $\NP$ language $L$, there exists a polynomial-time reduction $f$ mapping $L$ to 3CNF formulas such that:

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \frac{1}{2}
\end{align*}
\end{theorem}

@@

\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) For some constants $q \in \mathbb{N}$, the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard. To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ where $\Psi$ consists of $q$-ary predicates, such that

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \frac{1}{2}
\end{align*}
\end{theorem}

@@

\begin{lemma} (Equivalence between $\mathsf{Gap}$-$\mathsf{3SAT}$ and $\mathsf{Max}$-$\mathsf{qCSP}$). Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
\end{lemma}
%
\begin{proof}
To be written...
\end{proof}

>>>>

\begin{theorem} \label{pcpgapsat}($\PCP$ Theorem: $\mathsf{Gap3SAT}$-hardness)
The problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max3SAT}$ is $\NP$-hard. In other words there exists a constant $\alpha  \leq 1$ such that, for every $\NP$ language $L$, there exists a polynomial-time reduction $f$ mapping $L$ to 3CNF formulas such that:

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \alpha
\end{align*}
\end{theorem}

@@

\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) There exists constants $q \in \mathbb{N}$ and $\alpha \leq 1$ where the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard. To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ over domain $\Omega=\{0,1\}$ where $\Psi$ consists of $q$-ary predicates, such that

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \alpha
\end{align*}
\end{theorem}

@@

\begin{lemma} (Equivalnce between $\mathsf{GapMaxCSP}$and $\mathsf{GapMax3SAT}$ Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
\end{lemma}
%
\begin{proof}
Since any $\mathsf{3Sat}$ instance can be seen as a particular type of $\mathsf{3CSP}$ instance, one direction is immediate. Conversely, if we assume Theorem \ref{pcptheoremgapcsp}, we aim to find some constant $\alpha'$ such that there exist a reduction from an instance of $\mathsf{Gap}_{1-\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ for $q \in \mathbb{N}$ claimed in Theorem \ref{pcptheoremgapcsp} to an instance of $\mathsf{Gap}_{1-\alpha',1}\mathsf{Max3Sat}$.
%
Let the CSP $\Psi$ be comprised $q$-ary predicates $\psi$. For an instance $\I$ of $\Psi$, an constraint tuple $(S,\psi)$ can be expressed as a logical AND of $2^q$ clauses where each clause is a logical OR of $q$ variables or their negations. In other words, $(S,\psi)$ is essentially a CNF of width $q$ and of size at most $2^q$. We can then construct an ``equivalent" 3CNF as follows: add extra symbols $\Pi_{1}, \cdots, \Pi_{(q-3)2^{q}}$. It can be shown that there exists a 3CNF $\psi'_S$ of size at most $(q-2)2^{q}$ such that:

\begin{enumerate}
  \item For every $x \in \{0,1\}^q$ which causes $\psi(x) = 1$, there exists an assignment $\Pi$ such that $\psi'(x,\Pi) = 1$.
  \item Else if $\psi(x) = 0$, then for all assignments $\Pi$, $\psi'(x,\Pi) = 0$
\end{enumerate}

Now take the total 3CNF defined by a conjection of all such $\psi'_S$:
\[ \psi_{\I} = \bigwedge_{(S,\psi) \in \I}  \psi_S' \]

The total formula $\psi_\I$ is determined by at most $mq2^q$ number of clauses and at most $n + m(q-3)2^q$ number of variables.
%
If an instance $\I$ has all of its constraints satisfiable by some assignment, then by proprety (1) listed above, there must exist a assignment to the variables of the $\psi_{\I}$ such that it is satisifed.
%
On the other hand, if for all assignments to instance $\I$ only satisfy at most an $1 - \alpha$ fraction of constraints, then fraction of clauses satisfied can be at most $1 - \alpha + \alpha(1 - \frac{1}{(q-2)2^qa}) = 1 - \frac{\alpha}{(q-2)2^q}$. This is due to the fact that for each unsatisfied $\psi_S'$, the most number of its clauses which can be satisfied by any assignment will be $(q-2)2^q - 1$. Taking $\alpha' = \frac{\alpha}{(q-2)2^q}$
yields the required constant.
\end{proof}

### 2021-11-24 Complexity/UGC/uniquegames.tex
Grammar and additions to pcp section

\begin{definition}
  A problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction from $\GapDeltaULC$ to $P$.
\end{definition}

>>>>

\begin{definition}
  A promise problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction $f$ from $\GapDeltaULC$ to $P$, in the sense that for a $\GapDeltaULC$ instance $\I$:
  \begin{align*}
    \mathsf{Opt}(\I)  \geq 1 - \delta & \implies f(x) \in \mathsf{Yes} \\
    \mathsf{Opt}(\I)  < \delta & \implies f(x) \in \mathsf{No}
  \end{align*}
\end{definition}
